---
title: "Post-ARU Deployment Summary"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When you get the rest of the data from the Musselshell, finalize this datasheet and send it out - I'll add them to the OneDrive and the Dropbox
```{r Setup}
# Load in packages
packages <- c("mapview","tidyverse","janitor")
source(".\\R_Scripts\\Install_Load_Packages.R")
load_packages(packages)

```

```{r Read Data}
# read in the original metadata
metadat <- read_csv(".\\Data\\Metadata\\2023_ARUDeployment_Metadata_Orig123.csv") %>% clean_names

```

## Data cleaning:

- Check the unique values of each column

- Change lowercase miso/Miso/MISS to MISO, mush/Mush to MUSH, yell/Yell/YEll to YELL

Diagnostics on the data


```{r Data Cleaning}

# drop the first six columns
metadat <- metadat %>% select(-(1:6))

# replace the underscore with a dash in the point id 
metadat <- metadat %>% mutate(point_id = str_replace(point_id, "_","-"))

# separate point ID into point and ID
metadat <- metadat %>% separate(point_id, into = c("point","ID"), sep = "-")
# check the unique values of the point
unique(metadat$point)
# mutate to be correct values
metadat <- metadat %>% mutate(point = case_when(
  point == "Yell"~"YELL",
  point == "YEll"~"YELL",
  point == "YELl"~"YELL",
  point == "YELL"~"YELL",
  point == "MISS"~"MISO",
  point == "Miss"~"MISO",
  point == "MISO"~"MISO",
  point == "PRD"~"PRD",
  point == "ELI"~"ELI",
  point == "XSS"~"XSS",
  point == "AME"~"AME",
  point == "ISA"~"ISA",
  point == "MUSH"~"MUSH",
  point == "JDO"~"JDO",
  point == "YWM"~"YWM",
  point == "GMW"~"GMW",
  point == "CUL"~"CUL",
  point == "ROB"~"ROB",
  point == "SNO"~"SNO",
  point == "84"~"84",
  point == "82"~"82",
  point == "105"~"105",
  point == "104"~"104",
  point == "101"~"101",
  point == "103"~"103",
  point == "102"~"102",
  point == "97"~"97",
  point == "83"~"83",
  point == "203"~"203",
  point == "JUD"~"JUD"))

# combine again
metadat <- metadat %>% unite(point_id,point, ID, sep="-")

# Write this to a csv
write.csv(metadat,".\\Data\\Metadata\\2023_ARUDeployment_Metadata_6-8.csv")
```

# Summary of ARU Deployments
```{r Summary}
#Questions to answer:
#How many total ARUs deployed? 154

#How many ARUs in each river system?
arus_perriver <- metadat %>% group_by(river) %>% summarize(number_deployed = n())
# 97 along the Missouri, 17 out at the Musselshell, and 40 along the Yellowstone

#How many SMM vs AM?
arumodels_persurvey <- metadat %>% group_by(aru_model) %>% summarize(number_deployed = n())
# 105 AudioMoth
# 49 SongMeter Micro

#How many repeat monitoring points?
#How many habitat points?
#How many new points?
arus_persurvey <- metadat %>% group_by(site_use) %>% summarize(number_deployed = n())
# 88 habitat points
# 6 new points
# 60 repeat survey points

# How many strata for the habitat points?
metadat %>% filter(site_use == "Habitat Point") %>% group_by(landcover_strata) %>% summarize(number_deployed = n())
# wetland points: 34
# forest points: 25
# shrub/grassland points: 29


# figure out which were the upper missouri 
UMBEL <- metadat %>% filter(river == "Missouri River" & x <  -106.883799)



```



## Things that need editing

No navigation notes for some of them - need to make sure that if Daniel is not the one going back out for the points that others can find them. 


### Next steps

Vegetation surveys starting June 15th or as soon as the plants in your area are fully leafed out. 
Playback surveys starting June 15th and going until August 15th. 