---
title: "Cuckoo Survey Points 2023"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("data.table","sf","ggmap","terra","raster","mapview","tidyverse","rgdal","rgeos","XML","methods","FedData","rasterVis","tidyterra","spsurvey", "spData", "usmap","tmap","tmaptools","leaflet")
source(".\\R_Scripts\\Install_Load_Packages.R")
load_packages(packages)
```

# Study area
## Map of landcover data and buffer around rivers

```{r}
# Cadastrel data download http://ftpgeoinfo.msl.mt.gov/Data/Spatial/MSDI/Cadastral/Parcels/Wheatland/

# Testing out tmap with locs 
locs_dat <- fread(".\\Data\\Spatial_Data\\2022_ALLPoints.csv")
locs_dat <- na.omit(locs_dat)
#convert this into a spatial object
locs_sf <- locs_dat %>% 
  st_as_sf(coords=c("long", "lat")) %>% st_set_crs(32100)

hydro <- st_read(".\\Data\\Spatial_Data\\MT_Rivers\\hd43a.shp")
# Filter out only the rivers in our study area
names <- c("Missouri River","Yellowstone River","Musselshell River")
proj_hydro <- hydro %>% filter(NAME %in% names)
proj_hydro <- proj_hydro %>% dplyr::select(NAME, geometry)

#C:\Users\annak\OneDrive\Documents\UM\Research\Coding_Workspace\Cuckoo-Research\Data\Spatial_Data


```

Brief explanation of the buffer process and the landcover data used to buffer 
```{r}
# first, show entire land cover layer for the entire study site with the buff shapefile 
# read in the updated, reprojected raster
lcov <- terra::rast(".\\Data\\Spatial_Data\\NLCD_2019_MTLandcoverProjected.tiff")
lcov <- as.factor(lcov)

# # load in the legend 
legend <- pal_nlcd()
vals <- unique(lcov[[1]])
# # pull out the values of the legend that are in the values of my map
df <- legend[legend$ID %in% vals$label,]
levels(lcov) <- df[,c("ID","Class")]
# need to relate the colors that it plots the raster to specific variables 
colors=c("#5475A8","#FFFFFF","#E8D1D1","#E29E8C","#ff0000","#B50000","#D2CDC0","#85C77E","#38814E","#D4E7B0","#DCCA8F","#FDE9AA","#FBF65D","#CA9146","#C8E6F8","#64B3D5")


# read in the hydrology shapefile
hydro <- st_read("E:\\MT_Spatial_Data\\MT_Lakes_Streams\\hd43a\\hd43a.shp")
# Filter out only the rivers in our study area
names <- c("Missouri River","Yellowstone River","Musselshell River")
proj_hydro <- hydro %>% filter(NAME %in% names)
proj_hydro <- proj_hydro %>% dplyr::select(NAME, geometry)
# create a buffer around the hydro
buff <- st_buffer(proj_hydro, dist = 400)

```

Fill this in with the code from Figure 1 of the proposal 
```{r}
# plot for Figure 1 of proposal 
ggplot()+
  geom_spatraster(data = lcov_crop) + 
  scale_fill_manual(values=colors)+
  geom_sf(data=buff) + 
  theme(legend.position = "none") + 
  labs(title = "Study Area River Valleys", ylab = "Latitude", xlab = "Longitude") +
  theme_bw() + scale_y_continuous(limits = c(45.0,48.0),breaks = seq(45.0, 48.0, by = .5))



```

# Repeat monitoring points 
## Map of the long-term repeat monitoring points on top of the buffer layer and the landcover data

Add on the points to Figure 1
```{r}

```



```{r}
lcov <- terra::rast("C:\\Users\\annak\\OneDrive\\Documents\\UM\\Research\\Coding_Workspace\\Cuckoo-Research\\Data\\Spatial_Data\\NLCD_2019_MTLandcoverProjectedEdited.tiff")

lcov_mush <- terra::rast("C:\\Users\\annak\\OneDrive\\Documents\\UM\\Research\\Coding_Workspace\\Cuckoo-Research\\Data\\Spatial_Data\\NLCD_2019_MusselshellLandcover.tiff")

lcov_mush_masked <- terra::rast("D:\\MT_Spatial_Data\\For_GRTS\\\\NLCD_2019_MusselshellLandcover_MaskedToLidar.tiff")
#plot(lcov)

# try in plot first
tmap_mode("view")
# put landcover into for the habitat_stack - do lcov$Class
#tm_shape(lcov_mush_masked, raster.warp=FALSE) +  tm_raster()
#Error: Unable to transform stars

# plot for full land cover
#map <- tm_shape(lcov, raster.warp=FALSE) +  tm_raster()

# quick map
lcov_miso <- st_read("D:\\MT_Spatial_Data\\For_GRTS\\Landcover_Polygon_Musselshell_MaskedToRaster.shp")
qtm(lcov_miso, fill="Class") # how to have categorical fill?

# same thing:
tm_shape(lcov_miso)+#nothing on the map?
  tm_fill("Class")

tm_shape(locs_sf)+
  tm_dots(col = "organization", palette = "Dark2", n = 6)


# Use tmap options to set the basemap and overlay map permanently during the R session:
opts <- tmap_options(basemaps = c(Canvas = "Esri.WorldGrayCanvas", Imagery = "Esri.WorldImagery"),
    overlays = c(Labels = paste0("http://services.arcgisonline.com/arcgis/rest/services/Canvas/",
                                 "World_Light_Gray_Reference/MapServer/tile/{z}/{y}/{x}")))
tmap_options(opts)
tmap_mode("view")

# plot for survey locations
tm_basemap(leaflet::providers$Stamen.Watercolor)+
  tm_shape(locs_sf) + 
  tm_dots(col = "organization", palette = "Dark2", n = 6)+
  tm_tiles(server = 'https://stamen-tiles-{s}.a.ssl.fastly.net/terrain-background/{z}/{x}/{y}{r}.{ext}', group = NA, alpha = 1)
# this has a background but it is blank white

# try another one 
tm_basemap(server = 'http://maps.stamen.com/terrain/#8/47.001/-109.645', group = NA, alpha = 1)+
tm_shape(locs_sf) + 
  tm_dots(col = "organization", palette = "Dark2", n = 6)
# This works and will plot but doesn't have a basemap
# tm_shape(locs_sf) + 
#   tm_dots(col = "organization", palette = "Dark2", n = 6)

# Try another thing: https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html#basemaps-and-overlay-tile-maps
tm_basemap("Stamen.Watercolor")+
  tm_shape(locs_sf) + 
  tm_dots(col = "organization", palette = "Dark2", n = 6)
# Also only visualizing a blank background 

# try just plotting the river area of the rasters

```
# New Habitat Points
## Yellowstone GRTS plots

```{r Yellowstone Plot}
# read in lcov
lcov_yell_masked <- terra::rast("D:\\MT_Spatial_Data\\For_GRTS\\NLCD_2019_YellowstoneLandcover_MaskedToLidar.tiff")
# read in grts
yell_main <- fread("D:\\MT_Spatial_Data\\Yell_Main_Sample_MaskedToLidar.csv")
#convert this into a spatial object
yell_main_sf <- yell_main %>% 
  st_as_sf(coords=c("long", "lat")) %>% st_set_crs(32100)
# combine them in tmap


```

## Missouri GRTS points



## Musselshell GRTS points



