---
title: "Cuckoo Survey Points 2023"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("data.table","sf","ggmap","terra","raster","mapview","tidyverse","rgdal","XML","methods","FedData","rasterVis","tidyterra","spsurvey", "spData", "usmap")
source(".\\R_Scripts\\Install_Load_Packages.R")
load_packages(packages)
```

# Study area
## Map of landcover data and buffer around rivers

Brief explanation of the buffer process and the landcover data used to buffer 
```{r}
# first, show entire land cover layer for the entire study site with the buff shapefile 
# read in the updated, reprojected raster
lcov <- terra::rast(".\\Data\\Spatial_Data\\NLCD_2019_MTLandcoverProjected.tiff")
lcov <- as.factor(lcov)

# # load in the legend 
legend <- pal_nlcd()
vals <- unique(lcov[[1]])
# # pull out the values of the legend that are in the values of my map
df <- legend[legend$ID %in% vals$label,]
levels(lcov) <- df[,c("ID","Class")]
# need to relate the colors that it plots the raster to specific variables 
colors=c("#5475A8","#FFFFFF","#E8D1D1","#E29E8C","#ff0000","#B50000","#D2CDC0","#85C77E","#38814E","#D4E7B0","#DCCA8F","#FDE9AA","#FBF65D","#CA9146","#C8E6F8","#64B3D5")


# read in the hydrology shapefile
hydro <- st_read("E:\\MT_Spatial_Data\\MT_Lakes_Streams\\hd43a\\hd43a.shp")
# Filter out only the rivers in our study area
names <- c("Missouri River","Yellowstone River","Musselshell River")
proj_hydro <- hydro %>% filter(NAME %in% names)
proj_hydro <- proj_hydro %>% dplyr::select(NAME, geometry)
# create a buffer around the hydro
buff <- st_buffer(proj_hydro, dist = 400)

```

Fill this in with the code from Figure 1 of the proposal 
```{r}
# plot for Figure 1 of proposal 
ggplot()+
  geom_spatraster(data = lcov_crop) + 
  scale_fill_manual(values=colors)+
  geom_sf(data=buff) + 
  theme(legend.position = "none") + 
  labs(title = "Study Area River Valleys", ylab = "Latitude", xlab = "Longitude") +
  theme_bw() + scale_y_continuous(limits = c(45.0,48.0),breaks = seq(45.0, 48.0, by = .5))



```

# Repeat monitoring points 
## Map of the long-term repeat monitoring points on top of the buffer layer and the landcover data

Add on the points to Figure 1
```{r}

```

# New Habitat Points
## Yellowstone GRTS plots



## Missouri GRTS points



## Musselshell GRTS points



